Material Methods

### 2020-04-17 doc/MatMeth.Rmd
adding, improving methods section

### DMC1

```{r DMC1.section.outline, echo=FALSE, include=FALSE}
#cell certeria, number, t.test
#DMC1 -- leptotene and sygotene cells were identified by SC morphology (by degree of synapsis) (cite another paper?). foci counted manually, error ccalulated from a subset of cells counted twice (supplement)
```

<certeria for DMC1 stained cells (single cells) SC morphology used for staging.>
<subset of males from strains were used>

A subset of strains, *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were quantified for DMC1.  A total of `r length( (DMC1.DF %>% filter(stage == "Z"))$Original.Name)` and `r length( (DMC1.DF %>% filter(stage == "L"))$Original.Name)` of zygotene and leptotene cells were quantified respectively.

@@

# chromosome proportions

# sc bivalent used in models (aldready above)

# IFD / interference
```

The proportions of different chromosome classes (seperated by number of foci) was tested with chi-sqaure tests. <in R using the X package/function>

@@

**CURATION data ect** (i think these are still in results)

- The performance of the segmentaion algorithm was assessed by counting the number of objects segmented per single cell image (expected 19 or 20 in males and females respectively). All automated bivalent data was manually curated before analysis.

**chromosome proportion**

>>>>

### DMC1
```{r DMC1.section.outline, echo=FALSE, include=FALSE}
#cell certeria, number, t.test
#DMC1 -- leptotene and sygotene cells were identified by SC morphology (by degree of synapsis) (cite another paper?). foci counted manually, error ccalulated from a subset of cells counted twice (supplement)

#<A subset of strains, *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were quantified for DMC1.  A total of `r length( (DMC1.DF %>% filter(stage == "Z"))$Original.Name)` and `r length( (DMC1.DF %>% filter(stage == "L"))$Original.Name)` of zygotene and leptotene cells were quantified respectively.>
```
<certeria for DMC1 stained cells (single cells) SC morphology used for staging.>

The means of DMC1 foci per cell from a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were tested by t-tests.

@@

# chromosome proportions, number of CO per chrm, proportions perpared by chi-sq tests

#chrm chrteriztions -- SC length (chromatin compaction), foci position(raw and normalized), and interfocal distance, 
#these traits were fit into the main models (as mentioned above) ... t.tests were used to test between sexes and groups

# sc bivalent used in models (aldready above), the length of single SC compared by t-tests -- and fit into models
# Foci Position -- t-test  and put into models
# IFD / interference
```

@@

This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents. In this dataset, isolation rates per cell range from `r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 (molossinus^MSM^ male) to `r max(algorithm.stat.category$mean_pass.rate)`-- 0.72 (musculus^KAZ^ female).

**curation step**  From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` -- 10,458 bivalent objects were isolated by the image analysis software. After a human curation step ([@peterson2019]), `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` -- 9,829 single-bivalent observations remained.

To better understand the whole genome rec landscap -- we charecterizaed the number of crossover/MLH1 foci on single chromosomes (from the curated bivalent dataset).

We defined/classified each bivalent into a 'class' based on the number of MLH1 foci (range X to X). The **chromosome class proportions** different chromosome classes per strains by sex were tested with chi-sqaure tests with the prop.test function.

#main traits

chrm chrteriztions -- SC length (chromatin compaction), foci position(raw and normalized), and interfocal distance, 
these traits were fit into the main models (as mentioned above) ... t.tests were used to test between sexes and groups
